Cudele: A File System with Programmable Consistency and Durability

### 2017-03-20 paper/sections/0_abstract.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}
Today's large and highly-parallel workloads are bottlenecked by metadata
services because many processes end up accessing the same shared resoure. In
HPC, state-of-the-art file systems are abandoning POSIX because the
sychronization and serializiation overheads are too costly -- and sometimes
even unnceccessary -- for some of their applications.  While the performance
benefits are plain for these users, other applications that rely on stronger
consistency must be re-written or deployed on a different system.  We present
Cudele, a programmable file system that supports different degrees of
consistency and fault tolerance within the same namespace.  First, we take a
POSIX compliant file system and relax the consistency constraints by
implementing two metadata designs: delayed metadata update merge into the
global namespace and client local views of metadata updates.  Second, we store
the consistency and fault tolerance semantics in inodes so that subtrees within
the same namespace can be optimized for different workloads; the inodes are
programmable so that clients understand how to access the metadata in a
subtree.  Third, we present a theoretical framework for a metadata service and
a working implementation that provides the lowest common denominator needed to
implement a wide range of consistency/fault tolerance semantics that can be
benchmarked all on the same system.
\end{abstract}

>>>>

\begin{abstract}

HPC file systems are abandoning POSIX because the synchronization and
serialization overheads are too costly -- and often uneccessary -- for their
applications.  We provide an API for the client to (1) merge updates into the
global namespace and (2) assign policies to subtrees so that other clients
cannot interfere with the decoupled namespace.  This allows administrators to
optimize subtrees within the same namespace for different workloads. We draw
conclusions about the performance impact of previously unexplored
consistency/durability metadata designs and show that consistency can cause a
104\(\times\) slow down while merging updates (7\(\times\) slow down) and
maintaining durability (\(10\times\) slow down) have a more reasonable cost.

\end{abstract}

### 2017-03-20 paper/sections/1_introduction.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Today's client-server based file system metadata services have scalability
problems. It takes a lot of resources to server POSIX metadata requests and
applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. This is fine for small
workloads and file systems but as the system scales provisioning a metadata
server for every client is expensive and complicated.

Current hardware evolution and the rise of software-defined storage storage,
which uses techniqes like erasure coding, replication, and partitioning, have
ushered a new era of HPC computing; architectures are transitioning from
complex storage stacks with burst buffer, file system, object store, and tape
tiers to a two layer stack with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}. This trend exacerbates the metadata
scalability problem and has given rise to the serverless metadata services.
%\begin{listing}
\begin{minted}{python}
def hello():
\end{minted}
%\caption{JSON example} 
%\label{json-example}
%\end{listing}

HPC workloads are so metadata intensive that new management techniques are
advocating reducing synchronization and serialization overheads by transferring
these responsibilities to the client. We call this approach decoupling the
namespace and the semantics of consistency differ amongst systems.

@@

We propose subtree policies, an interface that lets future programmers control
how the system manages different parts of the namespace.  For performance one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  Our
system supports 3 forms of consistency and 2 forms of fault tolerance giving
the administrator a wide range of policies and optimizations depending on the
application's needs.

>>>>

% What is the problem?
Today's client-server based file system metadata services have scalability
problems. It takes a lot of resources to service POSIX metadata requests so
applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. Unfortunately,
provisioning a metadata server for every client is expensive and complicated.

% How does this fit in current landscape?
Current hardware evolution and the rise of software-defined storage storage,
which uses techniqes like erasure coding, replication, and partitioning, have
ushered in a new era of HPC computing; architectures are transitioning from
complex storage stacks with burst buffer, file system, object store, and tape
tiers to a two layer stack with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}. This trend exacerbates the metadata
scalability problem

% What is HPC doing?
To address this trend, HPC research has advocated client side processing and an
emphasis on serverless metadata services.  The techniques propose new metadata
management techniques that reduce synchronization and serialization overheads
by transferring these responsibilities to the client. We call this approach
``decoupling the namespace" and the semantics of consistency differ amongst
systems.  While the performance benefits are obvious for these users,
applications that rely on stronger consistency or durability guarantees must be
re-written or deployed on a different system.

@@

% What did we do
We propose subtree policies, an interface that lets future programmers control
how the system manages different parts of the namespace.  For performance one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  We
present Cudele, a prototype programmable file system that supports different
degrees of consistency and durability by exposing mechanisms used within the
file system as a client library.  Cudele supports 3 forms of consistency
(invisible, eventual, and strong) and 3 degrees of durability (none, local, and
global) giving the administrator a wide range of policies and optimizations
that can be custom fit to an application. Our contributions:

\begin{enumerate}

  \item a prototype that lets administrators program a range of
  consistency and durability semantics (9 permutations), allowing them to custom
  fit the storage system to the application.

  \item an API for assigning and programming policies to subtrees in the file
  system namespace.

  \item an analysis of recently proposed research systems compared against
  previously unexplored metadata designs.

\end{enumerate}

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improve
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). We also show the effect of having a metadata specific
file format in systems that are based on in-memory data structures.
Section~\ref{sec:related-work} places Cudele in the context of other related
work. Section~\ref{sec:posix-overheads} quantifies the cost of POSIX
consistency and system-defined durability and
Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
prototype and API. Section~\ref{sec:implementation} describes Cudele's
mechanisms and shows how re-using internal subsystems results in an
implementation of less than 500 lines of code. The evaluation in
Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
explored and previously unexplored metadata designs.
